Spin canonical divisors of log stacky curves

### 2015-07-01 stacky_write_up/spin_cring.tex
edited revised genus 2

The main result of this subsection is that for a log spin curve with no stacky points $(X,\Delta,L)$, the spin canonical ring $R_L$ is generated in degree at most $5$, with relations in degree at most $10$. The case that $\Delta = 0$ was completed in Reid \todo{cite Ried article}. The generation bound is shown in Lemma ~\ref{lem:generation-5} and the relations bound is shown in Lemma ~\ref{prop:relation-10}. We summarize the result in Corollary ~\ref{cor:g-2-presentation-bound}

The proofs of these results are similar to those in Neves ~\cite[Proposition III.4 and Proposition III.12]{neves:halfcan}, but slightly generalized as we assume $\Delta > 0$ instead of $\Delta = 0$ and do not assume there is a basepoint free pencil contained in $H^0(X,L).$

@@

\begin{lem}
\label{lem:generation-5}
Let $(X, \Delta, L)$ be a log spin curve of signature $(g;-;\delta)$ with $g \geq 2$ and $\Delta > 0.$ Let $s_1,s_2 \in H^0(X,2L)$ be two independent sections such that the vector subspace $V = ks_1 \oplus k s_2 \subseteq H^0(K)$ is basepoint free. Then, the map
\begin{align*}
	V \otimes H^0(nL) \rightarrow H^0((n+2)L)
\end{align*}
is surjective if $n \geq 5.$ In particular, $R_L$ is generated in degree at most $5.$
\end{lem}

@@

\begin{proof}[Proof of Lemma ~\ref{lem:generation-5}]
To show there are no new generators in degree at least $7$ if $\Delta = 0$ and in degree at least $6$ if $\Delta > 0$, it suffices to show that if $n \geq 5$ or if ($n = 4$ and $\Delta > 0$), the map
\begin{align*}
	H^0(nL) \otimes H^0(2L) \rightarrow H^0((n+2)L)
\end{align*}
is surjective. Indeed, since $V= ks_1 \oplus k s_2 \subseteq H^0(K)$ is basepoint free, by the basepoint-free pencil trick, (see \cite[Lemma 2.6]{saint-donat:proj} for a proof,) we obtain an exact sequence
$$\begin{tikzcd}
0 \ar r & H^0((n-2)L) \ar r & V \otimes H^0(nL) \ar{r}{f} & H^0((n+2)L)
\end{tikzcd}$$

We wish to show $f$ is surjective.
Note that $\dim_k \ker f = \dim_k H^0((n-2)L) = (n-3)(g - 1)$ using Riemann--Roch and the assumption $n \geq 5$ or ($n = 4$ and $\Delta > 0$). Note that here we are implicitly using the Remark ~\ref{rem:delta-not-1}.
Additionally, $\dim_k V \otimes H^0(nL) = 2 \cdot (n- 1)(g - 1)$, again using Riemann--Roch.
Therefore, $$\dim_k \im f = 2 \cdot (n- 1)(g - 1) -(n-3)(g - 1) = (n+ 1)(g - 1) = \dim_k H^0((n+2)L),$$
again using Riemann-Roch. Ergo, $f$ is surjective.
\end{proof}

@@

The next step is to bound the degrees of the relations of $R_L$. This is done in Proposition ~\ref{prop:relation-10}. The idea of the proof of Proposition ~\ref{prop:relation-10} is to use the basepoint-free pencil trick to show that if a relation lies in a sufficiently high degree, it lies in the ideal generated by the relations in lower degrees. We now fix notation for this ideal generated by lower degrees relations.

@@

Finally, we may then take $g = \sum_{i}^{}g_i x_i,h = \sum_{i}^{}h_i x_i,$ so that 
\begin{align*}
	f &\equiv \sum_{i}^{}a_i x_i \equiv \sum_{i}^{}(s_1g_i + s_2h_i)x_i \equiv s_1 \left( \sum_{i}^{}g_i x_i \right) + s_2 \left( \sum_{i}^{}h_i x_i \right) \\
	&\equiv s_1 g + s_2 h \bmod J_{L,d}.
\end{align*}
\end{proof}

\begin{prop}
\label{prop:relation-10}
Let $(X, \Delta, L)$ be a log spin curve of signature $(g;-;\delta)$ with $g \geq 2$ and $\Delta > 0.$ Then $I_L$ is generated in degree at most degree at most $10.$
\end{prop}
\begin{proof}
Suppose $f \in I_L$ with $\deg f \geq 11$.
To complete the proof, it suffices to show $f \in I_{L,d}'$. By Lemma ~\ref{lem:reducing-degree}, it suffices to show $s_1g+s_2h \in I_{L,d}'$ where $\phi(s_1),\phi(s_2) \in H^0(X,2L)$ are two sections so that $k\phi(s_1) \oplus k \phi(s_2) = V \subseteq H^0(K)$ is basepoint free. Consider the map
$$\begin{tikzcd}
V \otimes H^0((\deg f - 2)L) \ar {r}{f} & H^0((\deg f)L),
\end{tikzcd}$$
we know that $\phi(s_1)\phi(g) + \phi(s_2) \phi(h) \mapsto 0.$
So, by the explicit isomorphism given in the proof of the basepoint-free pencil trick, as shown in the proof of \cite{saint-donat:proj}[Lemma 2.6], there exists some $\rho \in k[x_1, \ldots, x_n]$ so that $\phi(\rho) \in H^0((\deg f - 4)L)$ satisfies $\phi(g) = \phi(s_2)\phi(\rho)$ and $\phi(h) = -\phi(s_1)\phi(\rho).$ Therefore, $g \equiv s_2 \rho \bmod I_{k,d-2}$ and $h \equiv -s_1 \rho \bmod I_{k,d-2}$. Hence,
\begin{align*}
	s_1g + s_2h \equiv s_1(s_2\rho) + s_2(-s_1 \rho) \equiv 0 \bmod J_{L,d}.
\end{align*}
\end{proof}

>>>>

The main result of this subsection is that for a log spin curve with no stacky points $(X,\Delta,L)$, the spin canonical ring $R_L$ is generated in degree at most $5$, with relations in degree at most $10$. The case that $\Delta = 0$ was completed in Reid \todo{cite Ried article}. For $\Delta > 0$, the generation bound is shown in Lemma ~\ref{lem:generation-5} and the relations bound is shown in Lemma ~\ref{prop:relation-10}. Thoughout this subsection, we will implicitly use Remark ~\ref{rem:delta-not-1}, which implies $\deg 2L \geq 2g$ and so $H^0(X, 2L)$ is basepoint free by Riemann-Roch. We summarize the results of this subsection in Corollary ~\ref{cor:g-2-presentation-bound}.

The proofs of this subsection results are similar to those in Neves ~\cite[Proposition III.4 and Proposition III.12]{neves:halfcan}. However, the statements are different, as we assume $\Delta > 0$ instead of $\Delta = 0$ and do not assume there is a basepoint free pencil contained in $H^0(X,L).$

@@

\begin{lem}
\label{lem:generation-5}
Let $(X, \Delta, L)$ be a log spin curve of signature $(g;-;\delta)$ with $g \geq 2$ and $\Delta > 0.$ Let $s_1,s_2 \in H^0(X,2L)$ be two independent sections such that the vector subspace $V = ks_1 \oplus k s_2 \subseteq H^0(K)$ is basepoint free. Then, the map
\begin{align*}
	V \otimes H^0(nL) \rightarrow H^0((n+2)L)
\end{align*}
is surjective if $n \geq 4.$ In particular, $R_L$ is generated in degree at most $5.$
\end{lem}

@@

\begin{proof}[Proof of Lemma ~\ref{lem:generation-5}]
To show there are no new generators in degree at least $6$, it suffices to show that if $n \geq 4$, the map
\begin{align*}
	H^0(nL) \otimes H^0(2L) \rightarrow H^0((n+2)L)
\end{align*}
is surjective. Indeed, since $V= ks_1 \oplus k s_2 \subseteq H^0(K)$ is basepoint free, by the basepoint-free pencil trick, (see \cite[Lemma 2.6]{saint-donat:proj} for a proof,) we obtain an exact sequence
$$\begin{tikzcd}
0 \ar r & H^0((n-2)L) \ar r & V \otimes H^0(nL) \ar{r}{f} & H^0((n+2)L)
\end{tikzcd}$$

We wish to show $f$ is surjective.
Note that $\dim_k \ker f = \dim_k H^0((n-2)L) = (n-3)(g - 1)$ using Riemann--Roch and the assumption $n \geq 4.$
Additionally, $\dim_k V \otimes H^0(nL) = 2 \cdot (n- 1)(g - 1)$, again using Riemann--Roch.
By rank-nullity, $$\dim_k \im f = 2 \cdot (n- 1)(g - 1) -(n-3)(g - 1) = (n+ 1)(g - 1) = \dim_k H^0((n+2)L).$$
Ergo, $f$ is surjective.
\end{proof}

@@

The next step is to bound the degrees of the relations of $R_L$ when $\Delta > 0$. This is done in Proposition ~\ref{prop:relation-10}. The idea of the proof of Proposition ~\ref{prop:relation-10} is to use the basepoint-free pencil trick to show that if a relation lies in a sufficiently high degree, it lies in the ideal generated by the relations in lower degrees. We now fix notation for this ideal generated by lower degrees relations.

@@

Finally, we may then take $g = \sum_{i}^{}g_i x_i, \: h = \sum_{i}^{}h_i x_i,$ so that 
\begin{align*}
	f &\equiv \sum_{i}^{}a_i x_i \equiv \sum_{i}^{}(s_1g_i + s_2h_i)x_i \equiv s_1 \left( \sum_{i}^{}g_i x_i \right) + s_2 \left( \sum_{i}^{}h_i x_i \right) \\
	&\equiv s_1 g + s_2 h \bmod J_{L,d}.
\end{align*}
\end{proof}

\begin{prop}
\label{prop:relation-10}
Let $(X, \Delta, L)$ be a log spin curve of signature $(g;-;\delta)$ with $g \geq 2$ and $\Delta > 0.$ Then $I_L$ is generated in degree at most degree at most $10.$
\end{prop}
\begin{proof}
Suppose $f \in I_L$ with $\deg f \geq 11$.
To complete the proof, it suffices to show $f \in I_{L,d}'$. By Lemma ~\ref{lem:reducing-degree}, it suffices to show $s_1g+s_2h \in I_{L,d}'$ where $\phi(s_1),\phi(s_2) \in H^0(X,2L)$ are two sections so that $k\phi(s_1) \oplus k \phi(s_2) = V \subseteq H^0(K)$ is basepoint free. Consider the map
$$\begin{tikzcd}
V \otimes H^0((\deg f - 2)L) \ar {r}{f} & H^0((\deg f)L),
\end{tikzcd}$$
we know that $\phi(s_1)\phi(g) + \phi(s_2) \phi(h) \mapsto 0.$
So, by the explicit isomorphism given in the proof of the basepoint-free pencil trick, as shown in the proof of \cite[Lemma 2.6]{saint-donat:proj}, there exists some $\rho \in k[x_1, \ldots, x_n]$ so that $\phi(\rho) \in H^0((\deg f - 4)L)$ satisfies $\phi(g) = \phi(s_2)\phi(\rho)$ and $\phi(h) = -\phi(s_1)\phi(\rho).$ Therefore, $g \equiv s_2 \rho \bmod I_{k,d-2}$ and $h \equiv -s_1 \rho \bmod I_{k,d-2}$. Hence,
\begin{align*}
	s_1g + s_2h \equiv s_1(s_2\rho) + s_2(-s_1 \rho) \equiv 0 \bmod J_{L,d}.
\end{align*}
\end{proof}
